Conservation decision-making under long transients

### 2019-08-13 slides/solarized.Rmd
WIP

@@

@@

@@

<img src="img/royalsociety/pensive-polarbear.jpg" width="400", style="float:center">

@@

```{r fpt}
passage %>% ggplot(aes(tau)) + 
  geom_histogram(binwidth = 200, col=txtcolor) + 
  geom_vline(aes(xintercept=det_tau), col=pal[1], lwd=2)
```

@@

p1 <- df %>% 
  ggplot((aes(value, fill=timeseries))) + 
  geom_histogram(alpha=0.9, show.legend = FALSE) + 
  geom_vline(aes(xintercept=0.27, col = "no switch"), col= pal[1], lwd=2) + 
  geom_vline(aes(xintercept=0.28, col = "no ghost"), col = pal[2], lwd=2)

@@

```{r}
samples3 <- read_csv("../data/posterior_ensemble.csv.xz")
samples3 %>%  
  ggplot() + 
  geom_histogram(aes(value), bins = 30, col=txtcolor)  +
  geom_vline(aes(xintercept = 0.023), col = pal[1], lwd = 2)

@@

# Descision Theory

@@

# Example management scenarios

@@

df <- mdp_planning(m_true$transition, m_true$reward, discount, model_prior = c(1), 
                   policy = no_policy, x0 = x0, Tmax = 100)

df %>% mutate(state = states[state], action = actions[action]) %>% 
  select(state, action, time) %>% 
  gather(series, state, -time) %>%
  ggplot(aes(time, state, col=series)) +
  geom_line(lwd=2) + 
  ylim(0,1.4)
```

@@

df %>% mutate(state = states[state], action = actions[action]) %>% 
select(state, action, time) %>% 
  gather(series, state, -time) %>%
  ggplot(aes(time, state, col=series)) +
  geom_line(lwd=2) + 
  ylim(0,1.4)
```

@@

## Adaptive managment: revising beliefs about the process

@@

### Prior beliefs

```{r priors}
prior <- dnorm(possible_a, 0.29, 0.01)
prior <- prior / sum(prior)
data.frame(a = possible_a, probability = prior) %>%
  ggplot(aes(a,prior)) + geom_bar(stat="identity", col=txtcolor) +
  geom_vline(aes(xintercept = true_a), lwd=1, lty=2, col=pal[1]) 
```

@@

### Learning too late

@@

### More agnostic prior beliefs

```{r priors_wide}
prior <- dnorm(possible_a, 0.3, 0.01)
prior <- prior / sum(prior)
data.frame(a = possible_a, probability = prior) %>%
  ggplot(aes(a,prior)) + geom_bar(stat="identity", col=txtcolor) +
  geom_vline(aes(xintercept = true_a), lwd=1, lty=2, col=pal[1]) 
```

@@

```{r eval=FALSE}
sim <- readRDS("sim-ghost-learning-2.Rds")

 sim$posterior %>% 
  data.frame(time = 1:Tmax) %>%
  filter(time %in% seq(1,Tmax, by = 5)) %>%
  gather(param, probability, -time, factor_key =TRUE) %>% 
  mutate(param = as.numeric(param)) %>% 
  ggplot(aes(param, probability, group = time, alpha = time)) + 
  geom_bar()

library(patchwork)
p1 + p2
```

@@

**Publishing Photo Competition**

>>>>

background-image: url(img/royalsociety/snake.jpg)
background-position: center
background-size: 120%
class: center, top

@@

]
.pull-right[
<img src="img/royalsociety/ballons.jpg" width="400", style="float:center">
]

@@

... Issues run deeper: transients hard to infer  *even under idealized conditions*

@@

#### *Stochastic* dynamics *qualitatively* alter the picture

]
.pull-right[
<img src="img/royalsociety/pensive-polarbear.jpg" width="400", style="float:center">
]

@@

```{r fpt}
passage %>% ggplot(aes(tau)) + 
  geom_histogram(binwidth = 200, fill=txtcolor) + 
  geom_vline(aes(xintercept=det_tau), col=pal[1], lwd=3)
```

@@

p1 <- df %>% 
  ggplot((aes(value, fill=timeseries))) + 
  geom_histogram(alpha=0.9, show.legend = FALSE) + 
  geom_vline(aes(xintercept=0.27, col = "no switch"), col= pal[1], lwd=3) + 
  geom_vline(aes(xintercept=0.28, col = "no ghost"), col = pal[2], lwd=3)

@@

```{r}
samples3 <- read_csv("../data/posterior_ensemble.csv.xz")
samples3 %>%  
  ggplot() + 
  geom_histogram(aes(value), bins = 30, fill=txtcolor)  +
  geom_vline(aes(xintercept = 0.023), col = pal[1], lwd = 3)

@@

# Decision Theory

@@

# Example Contexts

@@

df <- mdp_planning(m_true$transition, m_true$reward, discount, model_prior = c(1), 
                   policy = no_policy, x0 = x0, Tmax = 100)
```

```{r}
df %>% mutate(state = states[state], action = 5*actions[action]) %>% 
  select(state, action, time) %>% 
  gather(series, state, -time) %>%
  ggplot(aes(time, state, col=series)) +
  geom_line(lwd=2) + 
  ylim(0,1.3)
```

@@

df %>%
  mutate(state = states[state], action = 5*actions[action]) %>% 
  select(state, action, time) %>% 
  gather(series, state, -time) %>%
  ggplot(aes(time, state, col=series)) +
  geom_line(lwd=2) + 
  ylim(0,1.3)
```

@@

# Adaptive Management:

### Updating beliefs

<img src="img/royalsociety/waiting-in-the-shallows.jpg" width="500">

@@

## Prior beliefs

```{r priors}
prior <- dnorm(possible_a, 0.295, 0.005)
prior <- prior / sum(prior)
data.frame(a = possible_a, probability = prior) %>%
  ggplot(aes(a,prior)) + 
  geom_bar(stat="identity", fill=pal[1], alpha=0.6) +
  geom_vline(aes(xintercept = true_a), lwd=3, lty=2, col=pal[1]) + 
  ylim(0, 0.6)
```

@@

## Learning late

```{r}
sim <- readRDS("../notebook/sim-ghost-learning.Rds")
states <- seq(0,2, length=140)
actions <- states
Tmax <- 100

 sim$posterior %>% 
  data.frame(time = 1:Tmax) %>%
  filter(time %in% seq(1,Tmax, by = 10), time < 50) %>%
  gather(param, probability, -time, factor_key =TRUE) %>% 
  mutate(param = as.numeric(param)) %>% 
  ggplot(aes(x = param, y = probability, group = time, alpha = time)) + 
  geom_bar(stat="identity", position = "identity", show.legend = FALSE, fill=pal[1]) 
```

@@

## Outcome

```{r}
 sim$df %>% 
   mutate(state = states[state], action = 5*actions[action]) %>% 
  select(-value, -obs) %>% 
  gather(series, state, -time) %>% 
  ggplot(aes(time, state, color = series)) +
  geom_line(lwd=2) +
  ylim(0,1.3) +
  ylab("state / action")
```

## More agnostic prior beliefs

```{r priors_wide}
prior <- dnorm(possible_a, 0.3, 0.05)
prior <- prior / sum(prior)
data.frame(a = possible_a, probability = prior) %>%
  ggplot(aes(a,prior)) + geom_bar(stat="identity", fill=pal[1], alpha=0.5) +
  geom_vline(aes(xintercept = true_a), lwd=3, lty=2, col=pal[2]) +
  ylim(0,.3)
```

@@

```{r}
sim2 <- readRDS("../notebook/sim-ghost-learning-2.Rds")
states <- seq(0,2, length=200)
actions <- states
Tmax <- 100

 sim2$posterior %>% 
  data.frame(time = 1:Tmax) %>%
  filter(time %in% seq(1,Tmax, by = 10), time < 50) %>%
  gather(param, probability, -time, factor_key =TRUE) %>% 
  mutate(param = as.numeric(param)) %>% 
  ggplot(aes(x = param, y = probability, group = time, alpha = time)) + 
  geom_bar(stat="identity", position = "identity", show.legend = FALSE, fill=pal[1]) + 
   ylim(0,0.3)
```

```{r}
 sim2$df %>% 
   mutate(state = states[state], action = 5*actions[action]) %>% 
  select(-value, -obs) %>% 
  gather(series, state, -time) %>% 
  ggplot(aes(time, state, color = series)) +
  geom_line(lwd=2) + 
  ylim(0,1.3) +
  ylab("state / action")
```

# An emergent precautionary principle

<!-- Better to be ignorant than wrong -->

@@
